---
title: "vs_sup"
author: "drsimonturega"
date: "2025-12-16"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
```

# Virtual screening for high affinity guests for synthetic supramolecular receptors

### Orginal literature

<https://doi.org/10.1039/C5SC00534E>

### Function used in this analysis
I don't have a R data science package(library) yet... 
```{r df, echo=FALSE}

get_column_value <- function(d_frame, d_col) {
  # Use get() to retrieve the column by its string name
  get(d_col, d_frame)
}

hist_kde <- function(d_frame, d_col){
  #col = paste(d_frame,"$",d_col)
  #print(col)
  #print(class(col))
  #d_frame <- rlang::sym(d_frame)
  #dframe <- mutate
  #print(d_frame)
  #print(class(d_frame))
  #d_frame <- rlang::sym(d_col)
  #print(d_col)
  #print(class(d_col))
  data  = get(d_col, d_frame )
  hist(data, main=d_col, xlab = d_col, col="blue", freq = FALSE)
  dens <- density(data)
  lines(dens, lwd = 2, col = "black")
}
```

## Replication of our analysis in the manuscript using python


### We load and clean data from the supporting information

```{r}
df = read.csv("tab_gold_wt.csv", header = TRUE)
```

## Run some exploratory data analysis

### Data summary

```{r df}
summary(df)
```
 

### Colored Histogram with Different Number of Bins
```{r df}

hist_kde(df, "Part_buried")
```

```{r df, echo=FALSE}
hist(df$Part_buried, main="Part_buried", xlab = "Part_buried", col="blue", freq = FALSE)
dens <- density(df$Part_buried)
lines(dens, lwd = 2, col = "black")
```


## Including Plots

You can also embed plots, for example:

```{r df, echo=FALSE}
plot(df)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
