---
title: "vs_sup"
author: "drsimonturega"
date: "2025-12-16"
output:
  html_document: default
  pdf_document: default
---

```{r include=FALSE}
# Installing libraries required for our lecture
#installed rmarkdown ggplot2
#install.packages("reshape2")
library(reshape2)
library(ggplot2)
# You may have to instal tinytex
#tinytex::install_tinytex()
#install.packages("Metrics")
#install.packages("rtools")
#install.packages("randomForest")
library(randomForest)
library(Metrics)
library(caret)
#install.packages("neuralnet")
library(neuralnet)
```

# Virtual screening for high affinity guests for synthetic supramolecular receptors

### Orginal literature

<https://doi.org/10.1039/C5SC00534E>

### Functions used in this analysis
I don't have a R data science package(library) yet... 
```{r my_functions, echo=FALSE}

get_column_value <- function(d_frame, d_col) {
  # Use get() to retrieve the column by its string name
  # expand to use d_cal being a list
  get(d_col, d_frame)
}


hist_ylims <- function(dens){
  # Get y axis limits for a column from a data frame
  #data = get_column_value(d_frame, d_col)
  ylim = c(min(dens$y), c(max(dens$y)))
  return(ylim)
  }

int_made_even<- function(num){
  # make an int even
  if((num %% 2) != 0) 
    return(num+1)
  else
    return(num)
 }
  

hist_kde <- function(d_frame, d_col){
  # Colored Histogram with KDE line for clarity
  data  = get_column_value(d_frame, d_col )
  dens <- density(data)
  hist(data, main=d_col, xlab = d_col, ylim = (hist_ylims(dens)), col="blue", freq = FALSE)
  lines(dens, lwd = 2,  col = "black")
}

multi_plot <- function(d_frame,plot_t){
 # multi plot 
  n_col = ncol(d_frame)
  n_plot = int_made_even(n_col)
  col_names = colnames(d_frame)
  # make plots
  par(mfrow=c(2,(n_plot/2) ))
  for (col in col_names)
    plot_t(d_frame, col)
}

get_upper_tri <- function(mat){
  # split correlation matrix in have retain upper half
  mat[lower.tri(mat)]<- NA
  return(mat)
  }

get_lower_tri <- function(mat){
  # split correlation matrix in have retain upper half
  mat[upper.tri(mat)]<- NA
  return(mat)
}

reorder_cormat <- function(mat){
  # Use correlation between variables as distance
  dd <- as.dist((1-mat)/2)
  hc <- hclust(dd)
  mat <-mat[hc$order, hc$order]
  return(mat)
}

mat_rear <- function(mat){
  mat = round(cor(mat),2)
  mat = reorder_cormat(mat) # order Vars from low to high
  mat = get_upper_tri(mat)
  return(mat)
}

correl_mat_plot_2 <- function(mat, title){
  mat = round(cor(mat),2)
  mat = reorder_cormat(mat) # order Vars from low to high
  mat = get_upper_tri(mat)
  mat = melt(mat, na.rm = TRUE)
  ggplot(data = mat, aes(x = Var2, y = Var1, fill = value))+
    geom_tile(color = "white")+
    scale_fill_gradient2(low = "darkblue", high = "white", mid = "lightblue", 
      midpoint = 0, limit = c(-1,1), space = "Lab", 
      name="Pearson\nCorrelation") +
     theme_minimal()+
    # adds Pearson correlation values
    geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1,
      size = 12, hjust = 1)) +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
       panel.grid.major = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      axis.ticks = element_blank(),
      legend.justification = c(1, 0),
      legend.position = c(0.6, 0.7),
      legend.direction = "horizontal")+
      guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))+
    
    ggtitle(title) +
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_fixed()
}


```

## Replication of our analysis in the manuscript using R
We take a set df cleaned and prepared in the python verson of this lecture.

### We load cleaned data molecular descriptors from the supporting information
Students, visualise the molecular descriptors using R studio

```{r load datata science style df}
df = read.csv("tab_gold_wt.csv", header = TRUE)
```

### We load cleaned chemplp_score data from the supporting information

```{r load df for first manuscript plot}
df_chemplp_score = read.csv("chemplp_score.csv", header = TRUE)
dim(df_chemplp_score)
```

### Comparision of logKexp and ChemPLP_Score for training set from the manuscript
```{r logKexp and ChemPLP_Score linear regession}
corr_man = lm(ChemPLP_Score~logKexpt, data = df_chemplp_score)

summary(corr_man)$r.squared
```

```{r plotting logKexp vs ChemPLP_Score}
lm_rsq <- function(mod){
    m <- mod;
    rsq <- substitute(italic(r)^2~"="~r2, 
         list(r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(rsq));
}

ggplot(data = df_chemplp_score, aes(x=logKexpt, y=ChemPLP_Score)) +
  geom_point(size=3, shape=23) +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = 'black', 
                                    fill = NA, 
                                    size = 2),
    panel.background = element_blank()) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_text(x = 1, y = 60, label = lm_rsq(corr_man), parse = TRUE)
```

### Multivariate linear regression on the training set from the manuscript using eq. 2
The eq.2 includes molecular descriptors calculated in Gold. Then we slice our data set as in the manuscript and check dimensions.
```{r Slicing our df to replicate manuscript eq.2}
df_train <- df[1:54,]
df_test <- df[55:69,]
print("Dimensions of training set")
print(dim(df_train))
print("Dimensions of test set")
print(dim(df_test))
```
We run our regression model calculating a correct analysis of the model by plotting logKexp vs. logKcalc and calculating the rmse. This allows us to test our hypothesis, that we can use the molecular descriptors calculated in Gold to build our own predictive model for cage-guest binding and evaluate the model.
```{r eq.2 run regression analysis}
# no ligand_flexibility
eq2_lm = lm(logKexp~Ligand_clash + Ligand_torsion + Part_buried + Non.polar,data = df_train)
summary(eq2_lm)
```

```{r eq.2 predict logK for training set}
logKcalc = predict(eq2_lm, df_train)
```

```{r eq.2 calculate rmse for training set}
df_train["logKcalc"] = logKcalc
rmse_value <- rmse(df_train$logKcalc, df_train$logKexp)
```

```{r plot eq.2 training set logKexp vs logKcalc}
ggplot(data = df_train, aes(x=logKexp, y=logKcalc)) +
  geom_point(size=3, shape=23) +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = 'black', 
                                    fill = NA, 
                                    size = 2),
    panel.background = element_blank()) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_text(x = 1, y = -2, label = rmse_value, parse = TRUE)
paste("rmse = ", rmse_value)
```

### Multivariate linear regression on the training set from the manuscript using eq. 3
The eq.3 includes a molecular descriptor for ligand flexibility this was calculated outside of gold. Then we slice our data set as in the manuscript and check dimensions.
```{r Slicing our df to replicate manuscript eq.3}
df_train <- df[1:54,]
df_test <- df[55:69,]
print("Dimensions of training set")
print(dim(df_train))
print("Dimensions of test set")
print(dim(df_test))
```
Then run our regression model calculating a correct analysis of the model by plotting logKexp vs. logKcalc and calculating the rmse. This allows us to test our hypothesis, that we can use the molecular descriptors calculated in Gold to build our own predictive model for cage-guest binding and evaluate the model.
```{r run regression analysis eq.3}
# with ligand_flexibility
eq3_lm = lm(logKexp~Ligand_clash + Ligand_torsion + Part_buried + Non.polar
             + Ligand_flexibility, data = df_train)
summary(eq3_lm)
```

```{r predict logK for training set eq 3.}
logKcalc = predict(eq3_lm, df_train)
```


```{r calculate rmse for training set eq 3.}
df_train["logKcalc"] = logKcalc
rmse_value <- rmse(df_train$logKcalc, df_train$logKexp)
```

```{r training set eq 3. plot logKexp vs logKcalc}
ggplot(data = df_train, aes(x=logKexp, y=logKcalc)) +
  geom_point(size=3, shape=23) +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = 'black', 
                                    fill = NA, 
                                    size = 2),
    panel.background = element_blank()) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) 
paste("rmse = ", rmse_value)
```
### Analysis of new guests from the manuscripts
To rapidly test our model we computationaly screened the labs inventory using the eq. 3 model to quickly find molecules to test the predictive model with. 
```{r predict logK for test set eq.3}
logKcalc_test = predict(eq3_lm, df_test)
```

```{r calculate rmse for test set eq.3}
df_test["logKcalc"] = logKcalc_test
rmse_value <- rmse(df_test$logKcalc, df_test$logKexp)
```

```{r test set eq.3 plot logKexp vs logKcalc}
ggplot(data = df_test, aes(x=logKexp, y=logKcalc)) +
  geom_point(size=3, shape=23) +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = 'black', 
                                    fill = NA, 
                                    size = 2),
    panel.background = element_blank()) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) 
paste("rmse = ", rmse_value)

```

## END OF LECTURE ONE

## Modern data science approach using eq. 3

### Run some exploratory data analysis

#### Data summary

```{r lecture 2 df summary}
summary(df)
```
 we still have guest number the data frame we will remove it
 
```{r remove column with guest number}
 df <- df[-c(1)]
 summary(df)
```

#### Histogram with KDE line for clarity
The histogram is in blue bars. The Kernel Density Estimation line is and estimation of the histogram represented as a continuous line and is shown as a black line.
```{r Histogram with KDE line}
hist_kde(df, "Part_buried")
```

Its helpful to plot all dataframe columns
```{r multi histogram with KDE line}
multi_plot(df, hist_kde)
```

#### Correlation matrix heatmap
Over correlated molecular descriptor (GoldPLP functions)  columns can cause problems with our regression models. Too many correlated molecular descriptor cause a misrepresentation of those molecular descriptors in our regression model.
```{r Correlation matrix heatmap preparing the data}
cor_mat <- round(cor(df),2)
head(cor_mat)
```
Reshape our dataframe
```{r Correlation matrix heatmap reshape data}
melted_cor_mat <- melt(cor_mat)
head(melted_cor_mat)
```

```{r neater correlatio matrix_2}
correl_mat_plot_2(df, "Correlation plot for functions from GoldPLP")
```

You can make your correlation plot neater but they are not really pretty.

### Normalise our df 

```{r pre normalise}
head(df)
```

```{r normilise}
process <- preProcess(as.data.frame(df[1:5]), method=c("range"))
df[1:5] <- predict(process, as.data.frame(df[1:5]))
```

```{r post normilise}
head(df)
```
### A random forest decision tree method
#### Splitting our data in to training and test sets

``` {r random forest spliting our data}
set.seed(89)
sample <- sample.int(n = nrow(df), size = floor(.80*nrow(df)), replace = F)
train <- df[sample, ]
test  <- df[-sample, ]
print("Dimensions of training set")
print(dim(train))
print("Dimensions of test set")
print(dim(test))
```

#### Fit our regression model

Fit our regression model and produce a description of the model, be-aware this is not the analysis of our model.
```{r random forest fit model}
logK.rf <- randomForest(logKexp ~ ., data = train, mtry = 3, importance = TRUE)
print(logK.rf)
```

#### Plot logKexp vs logKcalc

Then run our regression model calculating a correct analysis of the model by plotting logKexp vs. logKcalc and calculating the rmse. This allows us to test our hypothesis, that we can use the molecular descriptors calculated in Gold to build our own predictive model for cage-guest binding and evaluate the model.
```{r random forest prediction of logK for training set}
train["logKcalc"] <- predict(logK.rf, train)
```

```{r random forest calaculate rmse traning set}
rmse_value <- rmse(train$logKcalc, train$logKexp)
```

```{r random forest traning set plot logKexp vs logKcalc}
ggplot(data = train, aes(x=logKexp, y=logKcalc)) +
  geom_point(size=3, shape=23) +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = 'black', 
                                    fill = NA, 
                                    size = 2),
    panel.background = element_blank()) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  geom_text(x = 1, y = -2, label = rmse_value, parse = TRUE)
  
paste("rmse = ", rmse_value)
```

```{r prediction of logK for random forest test set}
test["logKcalc"] <- predict(logK.rf, test)
```

```{r calaculate rmse random forest test set}
rmse_value <- rmse(test$logKcalc, test$logKexp)
```

```{r random forest test set plot logKexp vs logKcalc}
ggplot(data = test, aes(x=logKexp, y=logKcalc)) +
  geom_point(size=3, shape=23) +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = 'black', 
                                    fill = NA, 
                                    size = 2),
    panel.background = element_blank()) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) #+
  #geom_text(x = 1, y = -2, label = rmse_value, parse = TRUE)
  paste("rmse = ", rmse_value)
```
### A Multilayer Perceptron neural networks method
#### Splitting our data in to training and test sets
```{r MLP data split}
set.seed(89)
sample <- sample.int(n = nrow(df), size = floor(.80*nrow(df)), replace = F)
train <- df[sample, ]
test  <- df[-sample, ]
print("Dimensions of training set")
print(dim(train))
print("Dimensions of test set")
print(dim(test))
```

#### Fit our regression model

Then run our regression model calculating a correct analysis of the model by plotting logKexp vs. logKcalc and calculating the rmse. This allows us to test our hypothesis, that we can use the molecular descriptors calculated in Gold to build our own predictive model for cage-guest binding and evaluate the model.
```{r formulate and train the MLP}
# formulate eq 3 for our model
eq_3_form <- logKexp~Ligand_clash + Ligand_torsion + Part_buried + Non.polar + Ligand_flexibility
# Train the MLP model
model <- neuralnet(eq_3_form, data = train, hidden = c(5, 3), linear.output = TRUE, stepmax=1e7)
```

#### Plot logKexp vs logKcalc

```{r MLP prediction of logK for training set}
train["logKcalc"] <- predict(model, train)
```

```{r MLP training set calaculate rmse}
rmse_value <- rmse(train$logKcalc, train$logKexp)
```

```{r MLP training set plot logKexp vs logKcalc}
ggplot(data = train, aes(x=logKexp, y=logKcalc)) +
  geom_point(size=3, shape=23) +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = 'black', 
                                    fill = NA, 
                                    size = 2),
    panel.background = element_blank()) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) #+
  #geom_text(x = 1, y = -2, label = rmse_value, parse = TRUE)
  
paste("rmse = ", rmse_value)
```

```{r prediction of logK for MLP test set}
test["logKcalc"] <- predict(model, test)
#head(test)
```

```{r MLP test set calaculate rmse}
rmse_value <- rmse(test$logKcalc, test$logKexp)
```

```{r MLP test set plot logKexp vs logKcalc}
ggplot(data = test, aes(x=logKexp, y=logKcalc)) +
  geom_point(size=3, shape=23) +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = 'black', 
                                    fill = NA, 
                                    size = 2),
    panel.background = element_blank()) +
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) 
paste("rmse = ", rmse_value)
```
