---
title: "vs_sup"
author: "drsimonturega"
date: "2025-12-16"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
```

# Virtual screening for high affinity guests for synthetic supramolecular receptors

### Orginal literature

<https://doi.org/10.1039/C5SC00534E>

### Function used in this analysis
I don't have a R data science package(library) yet... 
```{r df, echo=FALSE}

get_column_value <- function(d_frame, d_col) {
  # Use get() to retrieve the column by its string name
  # expand to use d_cal being a list
  get(d_col, d_frame)
}


hist_ylims <- function(dens){
  # Get y axis limits for a column from a data frame
  #data = get_column_value(d_frame, d_col)
  ylim = c(min(dens$y), c(max(dens$y)))
  return(ylim)
  }


  

hist_kde <- function(d_frame, d_col){
  # Colored Histogram with KDE line for clarity
  data  = get_column_value(d_frame, d_col )
  dens <- density(data)
  hist(data, main=d_col, xlab = d_col, ylim = (hist_ylims(dens)), col="blue", freq = FALSE)
  lines(dens, lwd = 2,  col = "black")
}
```

## Replication of our analysis in the manuscript using python


### We load and clean data from the supporting information

```{r}
df = read.csv("tab_gold_wt.csv", header = TRUE)
```

## Run some exploratory data analysis

### Data summary

```{r df}
summary(df)
```
 we still have guest number the data frame we will remove it
 

 
```{r df}
 df <- df[-c(1)]
 summary(df)
```
```{r df}
col_names = colnames(df)
print(ncol(df))
print(col_names[1])
```

### Colored Histogram with KDE line for clarity
```{r df}

hist_kde(df, "Part_buried")
```

