---
title: "vs_sup"
author: "drsimonturega"
date: "2025-12-16"
output:
  html_document: default
  pdf_document: default
---

```{r include=FALSE}
library(tidyverse)
#tinytex::install_tinytex()
```

# Virtual screening for high affinity guests for synthetic supramolecular receptors

### Orginal literature

<https://doi.org/10.1039/C5SC00534E>

### Function used in this analysis
I don't have a R data science package(library) yet... 
```{r my_functions, echo=FALSE}

get_column_value <- function(d_frame, d_col) {
  # Use get() to retrieve the column by its string name
  # expand to use d_cal being a list
  get(d_col, d_frame)
}


hist_ylims <- function(dens){
  # Get y axis limits for a column from a data frame
  #data = get_column_value(d_frame, d_col)
  ylim = c(min(dens$y), c(max(dens$y)))
  return(ylim)
  }

int_made_even<- function(num){
  # make an int even
  if((num %% 2) != 0) 
    return(num+1)
  else
    return(num)
 }
  

hist_kde <- function(d_frame, d_col){
  # Colored Histogram with KDE line for clarity
  data  = get_column_value(d_frame, d_col )
  dens <- density(data)
  hist(data, main=d_col, xlab = d_col, ylim = (hist_ylims(dens)), col="blue", freq = FALSE)
  lines(dens, lwd = 2,  col = "black")
}

q_q_plot <- function(d_frame, d_col){
  # Generate a quartile quartile plot
  data  = get_column_value(d_frame, d_col )
  qqnorm(data, main = d_col,  pch = 1)
  qqline(data, col = "blue", lwd = 2)
}

multi_plot <- function(d_frame,plot_t){
 # multi plot 
  n_col = ncol(d_frame)
  print(n_col)
  n_plot = int_made_even(n_col)
  print(n_plot)
  col_names = colnames(d_frame)
  print(n_plot)
  # make plots
  par(mfrow=c(2,(n_plot/2) ))
  for (col in col_names)
    plot_t(d_frame, col)
 }

```

## Replication of our analysis in the manuscript using python


### We load cleaned data from the supporting information

```{r load datata science style df}
df = read.csv("tab_gold_wt.csv", header = TRUE)
```

## Run some exploratory data analysis

### Data summary

```{r }
summary(df)
```
 we still have guest number the data frame we will remove it
 

 
```{r}
 df <- df[-c(1)]
 summary(df)
```

```{r }
col_names = colnames(df)
print(ncol(df))
print(col_names[1])
int_made_even(3)
```

### Colored Histogram with KDE line for clarity
```{r }

hist_kde(df, "Part_buried")
```

```{r }
multi_plot(df, hist_kde)
```
### Quantile quantile plots
Our quantile quantile plots show us weather our data fits to a theoretical distribution 
```{r }
multi_plot(df, q_q_plot)

```
This is experimental data from 69 individual data points our team generated so we don't really expect it to fit a normal distribution. We would expect the replicates of **logKexp** to fit a normal distribution and the error quoted in the manuscript is at 95% confidence.
