---
title: "vs_sup"
author: "drsimonturega"
date: "2025-12-16"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
```

# Virtual screening for high affinity guests for synthetic supramolecular receptors

### Orginal literature

<https://doi.org/10.1039/C5SC00534E>

## Replication of our analysis in the manuscript using python

### We load and clean data from the supporting information

```{r}
df = read.csv("tab_gold_wt.csv", header = TRUE)
```

## Run some exploratory data analysis

### Data summary

```{r df}
summary(df)
```
 

### Colored Histogram with Different Number of Bins
```{r df, echo=FALSE}
hist(df$Part_buried, main="Part_buried", xlab = "Part_buried", col="blue", freq = FALSE)
dens <- density(df$Part_buried)
lines(dens, lwd = 2, col = "black")
```


## Including Plots

You can also embed plots, for example:

```{r df, echo=FALSE}
plot(df)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
